---
title: "02_Taxon_Assignment"
author: "Sarah Yerrace"
date: "2023-05-13"
output: html_document
---

This is the second part of the gut content metabarcoding pipeline after 01_with_cutadapt


# Load libraries

```{r}
library(dada2)
library(digest)
library(phyloseq)
library(tidyverse)
library(seqinr)
library(ape)
library(DECIPHER)
library(ade4)
library(filesstrings)
```

# Import the sequence-table

```{r}
seqtab.project.miseqrun1 <- as.matrix(
  read.delim(
    "../Output/Lionfish_sequence-table.tsv",
    header =TRUE,
    sep = "\t",
    row.names = 1
    )
  )
```

# Set up the Atlantic Fasfa file

Translate command to make | into ;
```{bash}
tr '|' ';' < ../Data/ALL_Atlantic_2_Feb_2017.fasta \
> ../Data/ALL_Atlantic_2_Feb_2017_edit.fasta
```

Translate commands to remove returns
```{bash}
tr '
' ';' < ../Data/ALL_Atlantic_2_Feb_2017_edit.fasta \
> ../Data/ALL_Atlantic_2_Feb_2017_edit2.fasta
```

Sed command to put returns where I want them
```{bash}
sed 's/>/\'$'\n/g' < ../Data/ALL_Atlantic_2_Feb_2017_edit2.fasta \
> ../Data/ALL_Atlantic_2_Feb_2017_edit3.fasta
```

sed command to put > back at beginning
```{bash}
sed 's/^/>/' < ../Data/ALL_Atlantic_2_Feb_2017_edit3.fasta \
> ../Data/ALL_Atlantic_2_Feb_2017_edit4.fasta
```

# yolo

```{r}
taxonomy <- assignTaxonomy(
  seqtab.project.miseqrun1,
  "../Data/ALL_Atlantic_2_Feb_2017_edit4.fasta",
  taxLevels = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "species"),
  tryRC = FALSE,
  minBoot = 50,
  outputBootstraps = TRUE,
  multithread = TRUE,
  verbose = TRUE
)


```


# Bash

```{bash}
cd ../Applications/
curl -O https://ftp.ncbi.nlm.nih.gov/blast/executables/blast+/LATEST/ncbi-blast-2.14.0+-x64-macosx.tar.gz
tar -xf ncbi-blast-2.14.0+-x64-macosx.tar.gz
```


Check if it's working

```{bash}
../Applications/ncbi-blast-2.13.0+/bin/blastx -h
```

# Make blast database

make the blast database

```{r, engine="bash"}
../Applications/ncbi-blast-2.13.0+/bin/makeblastdb \
-in ../Data/ALL_Atlantic_2_Feb_2017.fasta \
-dbtype nucl \
-out ../Output/Reference_DB
```

Write output to a fasta file that can be blasted
transpose table
```{r, eval=TRUE}
seqtab.nochim_trans <- as.data.frame(t(seqtab.nochim)) %>% rownames_to_column(var = "sequence") %>% 
    rowid_to_column(var = "OTUNumber") %>% mutate(OTUNumber = sprintf("otu%04d", 
    OTUNumber)) %>% mutate(sequence = str_replace_all(sequence, "(-|\\.)", ""))
```

convert to fasta file
```{r, eval=TRUE}
df <- seqtab.nochim_trans
seq_out <- Biostrings::DNAStringSet(df$sequence)

names(seq_out) <- str_c(df$OTUNumber, df$Supergroup, df$Division, df$Class, 
    df$Order, df$Family, df$Genus, df$Species, sep = "|")

Biostrings::writeXStringSet(seq_out, str_c( "Test_1_ASV.fasta"), compress = FALSE, 
    width = 20000)

#I had to move this file from my working directory to my data directory
```

examine fasta file
```{r, engine="bash", eval=TRUE}
head ../Data/Test_1_ASV.fasta
```

# Run Blast

change blastx to blastn
```{r, engine="bash"}
../Applications/ncbi-blast-2.13.0+/bin/blastn \
-query ../Data/Test_1_ASV.fasta \
-db ../Output/Reference_DB \
-out ../Output/Test_1_ASV.tab \
-num_threads 8 \
-max_target_seqs 1 \
-outfmt 6
```

Examine blast output
```{r, engine="bash"}
head -2 ../Output/Test_1_ASV.tab
wc -l ../Output/Test_1_ASV.tab
```
