---
title: "03_MARES_Taxon_Assignment"
author: "Sarah Yerrace"
date: "2023-05-13"
output: html_document
---

# Load libraries

```{r}
install.packages("data.table")
library(dada2)
library(digest)
library(phyloseq)
library(tidyverse)
library(seqinr)
library(ape)
library(DECIPHER)
library(ade4)
library(filesstrings)
library("data.table")
library(janitor)
```


# Import the sequence-table

```{r}
seqtab.project.miseqrun1 <- as.matrix(
  read.delim(
    "../Output/Lionfish_sequence-table.tsv",
    header =TRUE,
    sep = "\t",
    row.names = 1
    )
  )
```

# Bash

```{bash}
cd ../Applications/
curl -O https://ftp.ncbi.nlm.nih.gov/blast/executables/blast+/LATEST/ncbi-blast-2.14.0+-x64-macosx.tar.gz
tar -xf ncbi-blast-2.14.0+-x64-macosx.tar.gz
```


Check if it's working

```{bash}
../Applications/ncbi-blast-2.13.0+/bin/blastx -h
```

# Make blast database

make the blast database

```{r, engine="bash"}
../Applications/ncbi-blast-2.13.0+/bin/makeblastdb \
-in ../Data/MARES/MARES_NOBAR_16012023.fasta \
-dbtype nucl \
-out ../Output/Reference_DB_MARES_NOBAR
```

Write output to a fasta file that can be blasted
transpose table
```{r, eval=TRUE}
seqtab.nochim_trans <- as.data.frame(t(seqtab.nochim)) %>% rownames_to_column(var = "sequence") %>% 
    rowid_to_column(var = "OTUNumber") %>% mutate(OTUNumber = sprintf("otu%04d", 
    OTUNumber)) %>% mutate(sequence = str_replace_all(sequence, "(-|\\.)", ""))
```

convert to fasta file
```{r, eval=TRUE}
df <- seqtab.nochim_trans
seq_out <- Biostrings::DNAStringSet(df$sequence)

names(seq_out) <- str_c(df$OTUNumber, df$Supergroup, df$Division, df$Class, 
    df$Order, df$Family, df$Genus, df$Species, sep = "|")

Biostrings::writeXStringSet(seq_out, str_c( "Test_MARES_NOBAR_ASV.fasta"), compress = FALSE, 
    width = 20000)

#Move file from code to my data directory
```

examine fasta file
```{r, engine="bash", eval=TRUE}
head ../Data/Test_MARES_NOBAR_ASV.fasta
```

# Run Blast

change blastx to blastn
```{r, engine="bash"}
../Applications/ncbi-blast-2.13.0+/bin/blastn \
-query ../Data/Test_MARES_NOBAR_ASV.fasta \
-db ../Output/Reference_DB_MARES_NOBAR \
-out ../Output/Test_MARES_NOBAR_ASV.tab \
-num_threads 8 \
-max_target_seqs 1 \
-outfmt 6
```



Examine blast output
```{r, engine="bash"}
head ../Output/Test_MARES_NOBAR_ASV.tab
wc -l ../Output/Test_MARES_NOBAR_ASV.tab
```


# Filter Blast Hits and Merge Tables

```{r}
setwd("~/Yerrace_metabarcoding/Output")

#read ASV table into R
Test_ASV <- read.table("Test_MARES_NOBAR_ASV.tab")

#rename columns
colnames(Test_ASV) = c("OTUNumber", "Species", "pident", "length","mismatch","gapopen","qstart","qend","sstart","send","evalue","bitscore")

#merge tables by otu number
read_table<-left_join(Test_ASV, seqtab.nochim_trans, by = join_by("OTUNumber" == "OTUNumber"))
datatable(read_table, fillContainer = T)
```

Filter by sequence length

```{r}
# remove ASV with sequences shorter than 300 bp
read_table_sub <- subset(read_table, read_table$length>300) 

#remove ASV with pident <95%
read_table_sub2 <- subset(read_table_sub, read_table_sub$pident>98) 
```

Sum by Taxa
```{r}
#remove unneeded columns
read_table_summed<- read_table_sub2[ -c(1,3:13) ]

#remove prefix
read_table_summed$Species <- sub(".*__", "", read_table_summed$Species)

#summarize by species
by_species <- read_table_summed %>%
  group_by(Species)

read_table_species<-by_species %>%
  summarise_all(sum)

datatable(read_table_species, fillContainer = T)
```

Write table to CSV
```{r}
write.csv(read_table_species, "~/Yerrace_metabarcoding/Output/read_table.csv", row.names=FALSE)
```
